---
title: 'Observed data: initial check + summary'
author: "Annemarie"
date: "16/06/2020"
output: html_document
editor_options: 
  chunk_output_type: inline
---


# Load packages
```{r, include=FALSE}
library(dplyr)
library(tidyr)
library(reshape2)
library(stringr)
library(purrr)
library(EloRating)
library(domstruc) #install_github("danm0nster/domstruc")
library(ape)  # mantel test
library(vegan) # mantel test
library(einet) # effectiveness
library(igraph)
```

# Import data
```{r}
ANALYZE_aggCD <- read.csv(file="ANALYZE.aggDC.csv") %>% 
  rename(rowID=X) # only necessary if the first column is an X 
observed <- ANALYZE_aggCD 

head(observed)
str(observed)

# check 
unique(observed$actor)
unique(observed$subject)
unique(observed$date)

sum(observed$crowd)
sum(observed$displace)
sum(observed$total.DC)

```


## Helper functions & data
```{r}

#make matrix function (from dataframe, 1st col with row names)
matrix.please<-function(x) {
  m<-as.matrix(x[,-1])
  rownames(m)<-x[,1]
  m
}

#list of all valid color combinations/bird IDs
bird.list <- c("BBB","BBG","BBR","BGB","BGG","BRB","BRR","GBB","GBG","GGB","GGG","GGR","GRG","GRR","RBB","RBG","RBR","RGG","RGR","RRG","RRR")
length(bird.list)

## dataframe of all possible dyad combinations
dyad.list <- expand.grid(bird.list, bird.list) #head(dyad.list)
names(dyad.list) <- c("actor", "subject")
dyad.list <- subset(dyad.list, actor!=subject)
dyad.list$dyadID <- paste(dyad.list$actor, dyad.list$subject, sep="-")
dyad.list$actor <- as.character(dyad.list$actor)
dyad.list$subject <- as.character(dyad.list$subject)
head(dyad.list)

```

# Summarize data
Find total events per dyad (pool across days)

```{r}
observedXdyad <- observed
observedXdyad <- observedXdyad %>% 
                  dplyr::group_by(actor, subject) %>% 
                  dplyr::summarise(observed.crowd=sum(crowd),
                          observed.displace=sum(displace),
                          observed.totalCD=sum(total.DC)) 
observedXdyad <- as.data.frame(observedXdyad)   
observedXdyad$actor <- as.character(observedXdyad$actor)
observedXdyad$subject <- as.character(observedXdyad$subject)
```


## Check observed data 
Need to make sure that all birds in the group are included in each network (fill no interactions with zeros)

```{r}

# list of all dyads
head(dyad.list)
str(dyad.list)

# observed aggression by dyad by behavior
head(observedXdyad)
str(observedXdyad)

#subset(observedXdyad, actor==subject) #check that we don't have any self-loops

# merge data with full dyad list in case any didn't interact
observedXdyad.alldyads <- merge(observedXdyad, dyad.list, 
                           by=c("actor", "subject"), 
                           all.y=TRUE)
head(observedXdyad.alldyads)
str(observedXdyad.alldyads) #should have 420 dyads -> length(dyad.list$dyadID)

# convert NA to 0's
observedXdyad.alldyads[is.na(observedXdyad.alldyads)] <- 0

#check
head(observedXdyad.alldyads)
length(observedXdyad.alldyads$dyadID) # with 21 birds, we should have 420 dyads total


# cast into square matrix with actor vertically in first column, all subjects horizontally in first row
OBS.crd.mx <- dcast(observedXdyad.alldyads, actor~subject, value.var="observed.crowd")

# check that actors and subjects are correct
subset(observedXdyad.alldyads, actor=="BBB" & subject=="GGG")
subset(observedXdyad.alldyads, actor=="GGG" & subject=="BBB")
OBS.crd.mx[1,] #BBB crowds GGG
OBS.crd.mx[11,] # GGG crowds BBB
```


#### OBSERVED DATA summary
Summary of all measures for the observed crowds, displacements, and pooled data (totalCD)

initial check:  
  - Do the same or different dyads crowd and displace?
  - data sparsity
  - correlation
  
step 1 and step 3:
  - individual measures
  - network measures
  - dominance hierarchy measures
  - aggression strategy



```{r}
start.time <- Sys.time()

#### initial check: Summarize basic info
  
##  both dyads perform both behaviors? ----
  head(observedXdyad)
  length(observedXdyad$actor)


unique.crowds <- observedXdyad %>%
  filter(observed.crowd>0, observed.displace==0) # unique dyads that crowd
#View(unique.crowds)
length(unique.crowds$observed.crowd)

unique.displace <- observedXdyad %>%
  filter(observed.crowd==0, observed.displace>0) # unique dyads that displace
#View(unique.displace)
length(unique.displace$observed.displace)

dyads.CD <- observedXdyad %>%
  filter(observed.crowd>0, observed.displace>0) # dyads that perform both
#View(dyads.CD)
length(dyads.CD$actor)

head(observedXdyad.alldyads)
null.dyads <-  observedXdyad.alldyads %>%
  filter(observed.totalCD==0) # dyads that did not interact
#View(null.dyads)
length(null.dyads$actor)

# check 
n.dyads.CD <- length(unique.crowds$observed.crowd) + length(unique.displace$observed.displace) + length(dyads.CD$actor)
n.null.dyads<-length(null.dyads$actor)
n.dyads.CD+n.null.dyads # have to be the same
length(observedXdyad.alldyads$actor)

# combine
interaction <- c("crowd", "displace", "both", "none", "total")
n.dyads <- c(length(unique.crowds$observed.crowd), length(unique.displace$observed.displace),  length(dyads.CD$actor), length(null.dyads$actor), length(observedXdyad.alldyads$actor))
dyads.interacting<-data.frame(interaction, n.dyads)

 
## equal observations? ----
  basic.summary <- observedXdyad %>%
    summarize(ave.crowds = mean(observed.crowd), sd.crowds=sd(observed.crowd), 
              min.crowds=min(observed.crowd), max.crowds=max(observed.crowd),
              ave.displace = mean(observed.displace), sd.displace= 
              sd(observed.displace), min.displace=min(observed.displace), 
              max.displace=max(observed.displace), 
              ave.totalCD=mean(observed.totalCD), sd.totalCD=sd(observed.totalCD),
              min.total=min(observed.totalCD), max.total=max(observed.totalCD))

  n.Obcrowd <- sum(observedXdyad.alldyads$observed.crowd)
  n.Obdisplace <- sum(observedXdyad.alldyads$observed.displace)
  n.totalCD <- sum(observedXdyad.alldyads$observed.displace+ observedXdyad.alldyads$observed.crowd)
  
  prop.Obdisplace <- n.Obdisplace/(n.Obdisplace+n.Obcrowd)
  prop.Obcrowd <- n.Obcrowd/(n.Obdisplace+n.Obcrowd)

  obsdyads.crowd <- subset(observedXdyad.alldyads, observed.crowd>0)
  obsdyads.displace <- subset(observedXdyad.alldyads, observed.displace>0)

  n.Obcrowd.dyads <- length(obsdyads.crowd$actor)
  n.Obdisplace.dyads <- length(obsdyads.displace$actor)

  Obcrowd.density <- n.Obcrowd.dyads/length(dyad.list$dyadID)
  Obdisplace.density <- n.Obdisplace.dyads/length(dyad.list$dyadID)
  ObtotalCD.density <- (n.Obcrowd.dyads + n.Obdisplace.dyads)/length(dyad.list$dyadID)
  
 ## correlation ----
  
  ## pearson or spearman correlation (null dyads excluded)
# normal distribution?
hist(observedXdyad$observed.crowd)
hist(observedXdyad$observed.displace)
# linear relationship?
plot(observedXdyad$observed.crowd,observedXdyad$observed.displace) 

obs.cor <- cor(observedXdyad$observed.crowd, observedXdyad$observed.displace, 
               method = "spearman")

cor.test(observedXdyad$observed.crowd, 
         observedXdyad$observed.displace, 
         method = "spearman",
         exact = FALSE)

## correlation matrix  
# CROWD matrix
  obs.crowd.mx <- reshape2::dcast(observedXdyad.alldyads, actor~subject, value.var="observed.crowd") #head(ref.crowd.mx)
  obs.crowd.mx[is.na(obs.crowd.mx)] <- 0
  obs.crowd.mx <- matrix.please(obs.crowd.mx)
  
  # DISPLACEMENT matrix
  
  obs.disp.mx <- reshape2::dcast(observedXdyad.alldyads, actor~subject, value.var="observed.displace") #head(ref.disp.mx)
  obs.disp.mx[is.na(obs.disp.mx)] <- 0
  obs.disp.mx <- matrix.please(obs.disp.mx)
  
  # Mantel test
  #mantel.test(obs.crowd.mx, obs.disp.mx) # ape package
  obs.cor.result <- mantel(obs.crowd.mx, obs.disp.mx, method = "spearman") # vegan package
  obs.cor.matrix <- obs.cor.result$statistic

   # Pool basic summaries
  Obcd.basic <- cbind.data.frame(obs.cor, obs.cor.matrix,
                                n.Obcrowd, n.Obdisplace,
                                prop.Obcrowd, prop.Obdisplace, 
                                n.Obcrowd.dyads, n.Obdisplace.dyads,
                                Obcrowd.density, Obdisplace.density, 
                                n.totalCD,ObtotalCD.density
                                )
  
  ## START NETWORK ANALYSES ----
  
  ###### separate analyses by type (crowd vs displace)
  
  ### CROWD
  obs.crowd.mx <- reshape2::dcast(observedXdyad.alldyads, actor~subject, value.var="observed.crowd") #head(ref.crowd.mx)
  obs.crowd.mx[is.na(obs.crowd.mx)] <- 0
  obs.crowd.mx <- matrix.please(obs.crowd.mx)
  
  # network
   graph.obs.crowd <- graph_from_adjacency_matrix(obs.crowd.mx, "directed", weighted=T, diag=F)

  # individual-based metrics
   #out-degree
  Obcrowd.degree <- igraph::degree(graph.obs.crowd, mode = c("out"))
  
  # out-strength
  Obcrowd.strength <- strength(graph.obs.crowd, mode = c("out"), weights = E(graph.obs.crowd)$weight)
  
  # betweenness centrality
  Obcrowd.between <- igraph::betweenness(graph.obs.crowd, directed = TRUE,
                                         weights = E(graph.obs.crowd)$weight)
  
  
  # eigenvector centrality
  Obcrowd.eigen <- eigen_centrality(graph.obs.crowd, directed = TRUE, weights = E(graph.obs.crowd)$weight)
  Obcrowd.ec <- Obcrowd.eigen$vector
  Obcrowd.ec.value <- Obcrowd.eigen$value
  
# network-based metrics  
  # average path length
  Obcrowd.apl <- mean_distance(graph.obs.crowd) # average.path.length(graph.obs.crowd) 
  
  # efficiency
  Obcrowd.ei <- effective_information(graph.obs.crowd, effectiveness = FALSE)
  Obcrowd.eff <- Obcrowd.ei/log2(length(bird.list))
  
  #Find proportion unknown relationships, a measure of sparseness
  Obcrowd.prunk <- EloRating::prunk(obs.crowd.mx)
  Obcrowd.prunk.pu <- as.numeric(Obcrowd.prunk[1])
  Obcrowd.prunk.dyads <- as.numeric(Obcrowd.prunk[2]) 
  
  #Find linearity
  Oblin.crowd<- EloRating::h.index(obs.crowd.mx, loops = 1000)
  Obcrowd.linearity <- Oblin.crowd[3,2]

  #Find steepness
  Obcrowd.steepness <- steepness::getStp(obs.crowd.mx, method="Dij")
  
  #Triangle transitivity
  Obcrowd.transi <- EloRating::transitivity(obs.crowd.mx, runs = 1000)
  Obcrowd.transi.Pt <- Obcrowd.transi[1]  # proportion of transitive triads
  Obcrowd.transi.ttri <- Obcrowd.transi[2]  # triangle transitivity
  
# aggression strategy  
  # Compute focus & position
  Obcrowd.focus <- dom_focus(obs.crowd.mx)
  Obcrowd.position <- dom_position(obs.crowd.mx)
  Obcrowd.fp <- cbind.data.frame(Obcrowd.focus, Obcrowd.position)
  colnames(Obcrowd.fp) <- c("focus", "position")
  
  #Compute blur models
  Obcrowd.blur <- dom_make_blur_data(obs.crowd.mx)
  
  #Find strategy
  strategy.Obcrowd <- dom_categorize_strategy(data=Obcrowd.fp, blur_data=Obcrowd.blur)
  
  
  ### POOL crowd hierarchy
  Obcrowd.hier <- cbind.data.frame(Obcrowd.ec.value,
                                  Obcrowd.apl, 
                                  Obcrowd.ei, Obcrowd.eff, 
                                  Obcrowd.linearity, 
                                  Obcrowd.steepness, 
                                  Obcrowd.prunk.pu, Obcrowd.prunk.dyads,
                                  Obcrowd.transi.Pt, Obcrowd.transi.ttri,
                                  Obcrowd.focus, Obcrowd.position,
                                  strategy.Obcrowd
                                  )
  
  ### DISPLACEMENT
  
  obs.disp.mx <- reshape2::dcast(observedXdyad.alldyads, actor~subject, value.var="observed.displace") #head(ref.disp.mx)
  obs.disp.mx[is.na(obs.disp.mx)] <- 0
  obs.disp.mx <- matrix.please(obs.disp.mx)
  
  graph.obs.displace <- graph_from_adjacency_matrix(obs.disp.mx, "directed", weighted=T, diag=F)
  #plot(graph.obs.displace)

# individual-based      
  #out-degree
  Obdisp.degree <- igraph::degree(graph.obs.displace, mode = c("out"))

  # out- strength
  Obdisp.strength <- strength(graph.obs.displace, mode = c("out"), weights = E(graph.obs.displace)$weight)

   # betweenness centrality
   Obdisp.between <- igraph::betweenness(graph.obs.displace, directed = TRUE,
                                    weights = E(graph.obs.displace)$weight) 
  
  # eigenvector centrality
  Obdisp.eigen <- eigen_centrality(graph.obs.displace, directed = TRUE, weights = E(graph.obs.displace)$weight)
  Obdisp.ec <- Obdisp.eigen$vector
  Obdisp.ec.value <- Obdisp.eigen$value
   
# network-based     
  # average path length
  Obdisp.apl <- average.path.length(graph.obs.displace) # mean_distance(graph.obs.displace)
  
  # efficiency
  Obdisp.ei <- effective_information(graph.obs.displace, effectiveness = FALSE)
  Obdisp.eff <- Obdisp.ei/log2(length(bird.list))
  
  #Find proportion unknown relationships, a measure of sparseness
  Obdisp.prunk <- EloRating::prunk(obs.disp.mx)
  Obdisp.prunk.pu <- as.numeric(Obdisp.prunk[1])
  Obdisp.prunk.dyads <- as.numeric(Obdisp.prunk[2])
  
# dominance   
  #Find linearity
  Oblin.disp<- EloRating::h.index(obs.disp.mx, loops = 1000)
  Obdisp.linearity <- Oblin.disp[3,2]

  #Find steepness
  Obdisp.steepness <- steepness::getStp(obs.disp.mx, method="Dij")
  
  #Triangle transitivity
  Obdisp.transi <- EloRating::transitivity(obs.disp.mx, runs = 1000)
  Obdisp.transi.Pt <- Obdisp.transi[1]  
  Obdisp.transi.ttri <- Obdisp.transi[2]  

# aggression strategy    
  # Compute focus & position
  Obdisp.focus <- dom_focus(obs.disp.mx)
  Obdisp.position <- dom_position(obs.disp.mx)
  Obdisp.fp <- cbind.data.frame(Obdisp.focus, Obdisp.position)
  colnames(Obdisp.fp) <- c("focus", "position")
  
  #Compute blur models
  Obdisp.blur <- dom_make_blur_data(obs.disp.mx)
  
  #Find strategy
  strategy.Obdisp <- dom_categorize_strategy(data=Obdisp.fp, blur_data=Obdisp.blur)
  
  
  ### POOL disp hierarchy
  Obdisp.hier <- cbind.data.frame(
                                  Obdisp.ec.value,
                                  Obdisp.apl,  
                                  Obdisp.ei, Obdisp.eff,
                                  Obdisp.linearity, 
                                  Obdisp.steepness, 
                                  Obdisp.prunk.pu, Obdisp.prunk.dyads,
                                  Obdisp.transi.Pt, Obdisp.transi.ttri,
                                  Obdisp.focus, Obdisp.position,
                                  strategy.Obdisp
                                  )

  
  
  ### TOTAL (totalCD, all crowds + all displaces)
  
  obs.totalCD.mx <- reshape2::dcast(observedXdyad.alldyads, actor~subject, value.var="observed.totalCD") #head(ref.totalCD.mx)
  obs.totalCD.mx[is.na(obs.totalCD.mx)] <- 0
  obs.totalCD.mx <- matrix.please(obs.totalCD.mx)
  
  graph.obs.totalCD <- graph_from_adjacency_matrix(obs.totalCD.mx, "directed", weighted=T, diag=F)
  #plot(graph.obs.totalCD)
  
# individual-based metrics   
  #out-degree
  ObtotalCD.degree <- igraph::degree(graph.obs.totalCD, mode = c("out")) 
   
  # out- strength
  ObtotalCD.strength <- strength(graph.obs.totalCD, mode = c("out"), weights = E(graph.obs.totalCD)$weight)

  # betweenness centrality
  ObtotalCD.between <- igraph::betweenness(graph.obs.totalCD, directed = TRUE,
                                    weights = E(graph.obs.totalCD)$weight) 
  
  # eigenvector centrality
  ObtotalCD.eigen <- eigen_centrality(graph.obs.totalCD, directed = TRUE, weights = E(graph.obs.totalCD)$weight)
  ObtotalCD.ec <- ObtotalCD.eigen$vector
  ObtotalCD.ec.value <- ObtotalCD.eigen$value
   
# group-based metrics    
  # average path length
  ObtotalCD.apl <- average.path.length(graph.obs.totalCD) # mean_distance(graph.obs.totalCD)
  
  # efficiency
  ObtotalCD.ei <- effective_information(graph.obs.totalCD, effectiveness = FALSE)
  ObtotalCD.eff <- ObtotalCD.ei/log2(length(bird.list))  
 
  #Find proportion unknown relationships, a measure of sparseness
  ObtotalCD.prunk <- EloRating::prunk(obs.totalCD.mx)
  ObtotalCD.prunk.pu <- as.numeric(ObtotalCD.prunk[1])
  ObtotalCD.prunk.dyads <- as.numeric(ObtotalCD.prunk[2])
  
# dominance  
  #Find linearity
  Oblin.totalCD<- EloRating::h.index(obs.totalCD.mx, loops = 1000)
  ObtotalCD.linearity <- Oblin.totalCD[3,2]

  #Find steepness
  ObtotalCD.steepness <- steepness::getStp(obs.totalCD.mx, method="Dij")
  
  #Triangle transitivity
  ObtotalCD.transi <- EloRating::transitivity(obs.totalCD.mx, runs = 1000)
  ObtotalCD.transi.Pt <- ObtotalCD.transi[1]  # proportion of transitive triads
  ObtotalCD.transi.ttri <- ObtotalCD.transi[2]  # triangle transitivity

# aggression strategy    
  # Compute focus & position
  ObtotalCD.focus <- dom_focus(obs.totalCD.mx)
  ObtotalCD.position <- dom_position(obs.totalCD.mx)
  ObtotalCD.fp <- cbind.data.frame(ObtotalCD.focus, ObtotalCD.position)
  colnames(ObtotalCD.fp) <- c("focus", "position")
  
  #Compute blur models
  ObtotalCD.blur <- dom_make_blur_data(obs.totalCD.mx)
  
  #Find strategy
  Obstrategy.totalCD <- dom_categorize_strategy(data=ObtotalCD.fp, blur_data=ObtotalCD.blur)
  
  
  ### POOL totalCD hierarchy
  ObtotalCD.hier <- cbind.data.frame(
                                  ObtotalCD.ec.value,
                                  ObtotalCD.apl,
                                  ObtotalCD.ei, ObtotalCD.eff,
                                  ObtotalCD.linearity, 
                                  ObtotalCD.steepness, 
                                  ObtotalCD.prunk.pu, ObtotalCD.prunk.dyads,
                                  ObtotalCD.transi.Pt, ObtotalCD.transi.ttri,
                                  ObtotalCD.focus, ObtotalCD.position,
                                  Obstrategy.totalCD
                                  )
  
  
  ## correlation individual-based metric
  
  # observed correlation out-strength
#hist(Obcrowd.strength) # normal distribution?
#hist(observed$Obdisp.strength)
plot(Obcrowd.strength,Obdisp.strength)  # linear relationship?

cor.test(observed.ID$Obcrowd.strength, 
         observed.ID$Obdisp.strength, 
         method = "spearman",
         exact = FALSE)
  obs.cor.strength <- cor(Obcrowd.strength, Obdisp.strength,  
               method = "spearman")
  
  ## observed correlation betweenness
#hist(observed$Obcrowd.between) # normal distribution?
#hist(observed$Obdisp.between)
plot(Obcrowd.between,Obdisp.between)  # linear relationship?

cor.test(observed.ID$Obcrowd.between, 
         observed.ID$Obdisp.between, 
         method = "spearman",
         exact = FALSE)

  obs.cor.between <- cor(Obcrowd.between, Obdisp.between,  
               method = "spearman")
  
  # observed correlation eigenvector
#hist(observed$Obcrowd.ec) # normal distribution?
#hist(observed$Obdisp.ec) # linear relationship?
plot(Obcrowd.ec,Obdisp.ec) 

cor.test(observed.ID$Obdisp.ec, 
         observed.ID$Obcrowd.ec, 
         method = "spearman",
         exact = FALSE)  

obs.cor.eigen <- cor(Obcrowd.ec, Obdisp.ec,  
               method = "spearman")
  
# POOL ALL DATA
observed.summary <- cbind.data.frame(Obcd.basic, Obcrowd.hier, Obdisp.hier, ObtotalCD.hier,obs.cor.strength, obs.cor.between, obs.cor.eigen)

observed.summary.individual<-cbind.data.frame(Obcrowd.degree, 
                                              Obcrowd.strength,  
                                              Obcrowd.between, 
                                              Obcrowd.ec, 
                                              Obdisp.degree, 
                                              Obdisp.strength, 
                                              Obdisp.between,
                                              Obdisp.ec,
                                              ObtotalCD.degree,
                                              ObtotalCD.strength,
                                              ObtotalCD.between,
                                              ObtotalCD.ec)

# Check time
end.time <- Sys.time()

#Time to run:
end.time - start.time


#check the data
t(observed.summary)
t(observed.summary.individual)

# Write data
write.csv(observed.summary, file = "ANALYZE-SUMM.observed.csv")
write.csv(observed.summary.individual, file = "ANALYZE-SUMM.observed.individual.csv")
```