---
title: 'Reference model 2: Sparse data'
author: "van der Marel, Hobson, Prasher"
date: "15/06/2020"
output: html_document
---

Here, we perform all the analyses for reference model 2

# Load packages
```{r, include=FALSE}
library(dplyr)
library(tidyr)
library(reshape2)
library(stringr)
library(purrr)
library(EloRating)
library(domstruc) #install_github("danm0nster/domstruc")
library(vegan) # mantel test
library(einet) # effectiveness
library(igraph)
```

# Import data
```{r}
ANALYZE_aggCD <- read.csv(file="ANALYZE.aggDC.csv") %>% 
  rename(rowID=X) # only necessary if the first column is an X 
observed <- ANALYZE_aggCD 

head(observed)
str(observed)

# check 
unique(observed$actor)
unique(observed$subject)
unique(observed$date)

sum(observed$crowd)
sum(observed$displace)
sum(observed$total.DC)

```
# Summarize observed data
Find total events per dyad (pool across days)

```{r}
observedXdyad <- observed
observedXdyad <- observedXdyad %>% 
                  dplyr::group_by(actor, subject) %>% 
                  dplyr::summarise(observed.crowd=sum(crowd),
                          observed.displace=sum(displace),
                          observed.totalCD=sum(total.DC), 
                          .groups = 'drop') 
observedXdyad <- as.data.frame(observedXdyad)   
observedXdyad$actor <- as.character(observedXdyad$actor)
observedXdyad$subject <- as.character(observedXdyad$subject)
```



## Helper functions & data
```{r}

#make matrix function (from dataframe, 1st col with row names)
matrix.please<-function(x) {
  m<-as.matrix(x[,-1])
  rownames(m)<-x[,1]
  m
}

#list of all valid color combinations/bird IDs
bird.list <- c("BBB","BBG","BBR","BGB","BGG","BRB","BRR","GBB","GBG","GGB","GGG","GGR","GRG","GRR","RBB","RBG","RBR","RGG","RGR","RRG","RRR")
length(bird.list)

## dataframe of all possible dyad combinations
dyad.list <- expand.grid(bird.list, bird.list) #head(dyad.list)
names(dyad.list) <- c("actor", "subject")
dyad.list <- subset(dyad.list, actor!=subject)
dyad.list$dyadID <- paste(dyad.list$actor, dyad.list$subject, sep="-")
dyad.list$actor <- as.character(dyad.list$actor)
dyad.list$subject <- as.character(dyad.list$subject)
head(dyad.list)

```


#### 1. Create reference model

The goal of this reference model is to assess how sparser displacement data may affect our summary statistics. To do this we subsampled our displacement event data so that the number of displacements was reduced to equal the number of crowds (which were the rarer behavior type in our observations).

There are 2 parts to this model:
Part A: subsample displacements to equal the rare interaction type and add observed crowds to this dataframe

Part B: randomly re-allocate the total observed crowds and subsample displacements back to the 2 interaction types. 
```{r}
## get data ready

# displacements in long format (1 event at a time)
obs.disp.long <- select(observed, actor, subject, displace) %>%
  slice(rep(seq_len(n()), displace)) %>% 
  select(-displace) %>%
  mutate(disp.event = 1)
sum(obs.disp.long$disp.event)

# total number of crowds
n.obs.crowd<-sum(observed$crowd)  


replicates <- 100

run=2

# make empty dataframe to write loop results into
ref.model <- data.frame(runID=character(),
                        actor=character(),
                       subject=character(),
                       random.crowd=numeric(),
                       random.displace=numeric()
                       )
 
##  loop

for (run in 1:replicates) {
  r.seed <- run
  RNGkind(sample.kind="default") 
  set.seed(r.seed)
  
  ### PART A: SUBSAMPLE DISPLACEMENTS ####
  sample.disp <- sample_n(obs.disp.long, size = n.obs.crowd, replace = FALSE)
 
  ## summarize displacements by dyad
  # finds n disps per actor by date for trimmed data
  sample.dispXday <- sample.disp %>% 
    group_by(actor, subject) %>% 
    summarise(sample.displace=n(), .groups = 'drop')  %>%
    # mutate(behavior="displace") %>% 
    ungroup()
    sample.dispXday<-as.data.frame(sample.dispXday)
  
  
  # check whether displacements add up to number of crowds per run
  check.displace <- group_by(sample.dispXday) %>%
    summarise(n.displace=sum(sample.displace), length(sample.displace), .group='drop')
  head(check.displace)
  
  ### ADD OBSERVED CROWDS
  # keep only crowds from raw data
  obs.crowd <- select(observedXdyad, -observed.displace, -observed.totalCD) 
  total.obs.crowd<-sum(obs.crowd$crowd) # find total number of crowd events
  
 
  ### COMBINE observed crowds and subsampled displacements per runID
   runID <- rep(paste0("run",str_pad(run, 3, side="left", pad = "0")), nrow(obs.crowd)) 
  
  run.data<-cbind.data.frame(runID, obs.crowd)
   
  run.data <- full_join(run.data, 
                            sample.dispXday, by=c("actor", "subject")) %>%
    select(runID, everything()) %>%
    replace_na(list(observed.crowd = 0, sample.displace = 0)) %>%
    mutate(total.DC=sample.displace + observed.crowd) %>%
    arrange(runID)
#head(run.data)
#str(run.data)
    
  #### PART B: RE-ALLOCATE ####
  #re-set sampling conditions (use same seed as above, using loop to set r.seed)
  RNGkind(sample.kind="default") 
  set.seed(r.seed)
  
  ref.data <-  run.data %>%
                rowwise() %>%
                dplyr::mutate(random.displace = sample(0:total.DC, 1),
                              random.crowd = total.DC-random.displace)
  ref.data <- as.data.frame(ref.data)
  ref.model <- rbind.data.frame(ref.model, ref.data)
  
  }
  
# export file
write.csv(ref.model, file = "ANALYZE_refmodel2.csv")
  
```

check whether everything ran properly
```{r}
## check ref model output

#check that the randomized interactions vary from the observed interactions
ref.model$random.displace == ref.model$displace
ref.model$random.crowd == ref.model$crowd
#check that runs are different from each other
head(subset(ref.model, runID=="run001"))
head(subset(ref.model, runID=="run002"))
head(subset(ref.model, runID=="run003"))
head(subset(ref.model, runID=="run011"))
#check that all runs ran as expected
unique(ref.model$runID)
length(unique(ref.model$runID))

## loop to compare random to observed displacements for each runID
uniqrun = unique(ref.model$runID)
randvobs = data.frame(nmatches = numeric(), compar = character())
nrowxrun = nrow(observedXdyad)

for(i in uniqrun){
        nmatches = sum(ref.model[ref.model$runID==i,]$random.displace==ref.model[ref.model$runID==i,]$sample.displace)
        run = i
        df = as.data.frame(nmatches, run)
        randvobs=rbind.data.frame(randvobs,df)
        
        if(sum(randvobs[,1]==nrowxrun)>0){
                                break
                        } #if there is at least 1 occurrence of a complete match between random.displace and observed.displace columns, then break the loop. No more comparisons will be made. Look at end of randvobs to see which runs had a complete match
}

#check
randvobs


#loop that compares random.crowds (or sample.displace) from each run to all other runs 
uniqrun = unique(ref.model$runID)
empdf = data.frame(nmatches = numeric(), compar = character())
nrowxrun = nrow(observedXdyad)

        for (i in uniqrun) {
                
                #make uniqrunj (a list of all runIDs that come after i)
                n= which(!is.na(match(uniqrun,i))) 
                uniqrunj = uniqrun[n:length(uniqrun)] 
                uniqrunj = uniqrunj[uniqrunj!=i] 
                        
                for (j in uniqrunj) {
                  
                        nmatches = sum(ref.model[ref.model$runID==i,]$sample.displace==ref.model[ref.model$runID==j,]$sample.displace)    # gives the number of dyads for which the number of crowds in run i matches that in run j
                        
                        compar = paste(i,j, sep="-") #a vector to keep track of which runs are being compared
                        df = as.data.frame(nmatches,compar)
                        empdf=rbind.data.frame(empdf,df)
                  
                        if(sum(empdf[,1]==nrowxrun)>0){
                                break
                        } #if there is at least 1 occurrence of a complete match between random.crowd columns from the two runs, then break the inner loop
                        
                }
        
                val= as.numeric(substr(i,4,6))
                if(val %% 10==0) {
              cat(paste0("iteration: ", val, "\n"))
                   }# see loop progress at every 10 iterations
                               
                if(sum(empdf[,1]==nrowxrun)>0){
                        break
                } #if there is at least 1 occurrence of a complete match between random.crowd columns from the two runs, then break the entire loop - no more comparisons will be made. Look at end of empdf to see which runs had a complete match
                
        } #returns empdf with number of matches for each pair of runs being compared
tail(empdf)

```




#### 2. summarize ref model
The summary results of the observed data are calculated in the 'observed data.rmd' file and saved in "ANALYZE-SUMM.observed.csv"


Now, we can summarize our results using reference model 2.

```{r}
ref.model <- ref.model.sparse %>% select(runID, everything()) %>%
  rename(observed.crowd=obs.dyad.crowd, sample.displace=disp.events.subsampled, 
         total.DC=total.agg.REF2, random.displace=subsamp_realloc_displace, 
         random.crowd=realloc_crowd)

# or import from your working directory
refmodel2<- read.csv("ANALYZE_refmodel2.csv")
ref.model<-refmodel2

```


# Loop to produce all reference model summaries (except strategies)

```{r}

#make empty dataframe to fill
ref.model.summaries <- data.frame(run.code=character(), 
                                  run.cor=numeric(), 
                                  n.Rcrowd=numeric(), 
                                  n.Rdisplace=numeric(),
                                  prop.Rcrowd=numeric(), 
                                  prop.Rdisplace=numeric(), 
                                  n.Rcrowd.dyads=numeric(), 
                                  n.Rdisplace.dyads=numeric(),
                                  Rcrowd.degree=numeric(),
                                  Rcrowd.strength=numeric(),
                                  Rcrowd.between=numeric(),
                                  Rcrowd.ec=numeric(), 
                                  Rcrowd.ec.value=numeric(),
                                  Rcrowd.apl=numeric(),
                                  Rcrowd.ei=numeric(),
                                  Rcrowd.eff=numeric(),
                                  Rcrowd.linearity=numeric(), 
                                  Rcrowd.steepness=numeric(), 
                                  Rcrowd.transi.Pt=numeric(), 
                                  Rcrowd.transi.ttri=numeric(),
                                  Rdisp.degree=numeric(),
                                  Rdisp.strength=numeric(),
                                  Rdisp.between=numeric(),
                                  Rdisp.ec=numeric(), 
                                  Rdisp.ec.value=numeric(),
                                  Rdisp.linearity=numeric(), 
                                  Rdisp.steepness=numeric(), 
                                  Rdisp.apl=numeric(),
                                  Rdisp.ei=numeric(),
                                  Rdisp.eff=numeric(),
                                  Rdisp.transi.Pt=numeric(), 
                                  Rdisp.transi.ttri=numeric(),
                                  Rcrowd.density=numeric(), 
                                  Rdisplace.density=numeric(),
                                  Rcrowd.prunk.pu=numeric(), 
                                  Rcrowd.prunk.dyads=numeric(),
                                  Rdisp.prunk.pu=numeric(), 
                                  Rdisp.prunk.dyads=numeric(),
                                  R.cor.strength=numeric(),
                                  R.cor.between=numeric(),
                                  R.cor.eigen=numeric()
                                  )

run.s <- as.character(unique(ref.model$runID))
n.total.dyads <- length(dyad.list$dyadID) #total number of dyads 

start.time <- Sys.time()

for(run in seq_along(unique(ref.model$runID))){
  run.code <- run.s[run]
  print(run.code)
  run.data <- subset(ref.model, runID==run.code)
  
  #run.corr <- subset(COR.refmodel, runID==run.code)
  run.cor <- cor(run.data$random.displace, run.data$random.crowd, method = "spearman")
  
  #summarize basic info
  n.Rcrowd <- sum(run.data$random.crowd)
  n.Rdisplace <- sum(run.data$random.displace)
  prop.Rdisplace <- n.Rdisplace/(n.Rdisplace+n.Rcrowd)
  prop.Rcrowd <- n.Rcrowd/(n.Rdisplace+n.Rcrowd)

  dyads.crowd <- subset(run.data, random.crowd>0)
  dyads.displace <- subset(run.data, random.displace>0)

  n.Rcrowd.dyads <- length(dyads.crowd$actor)
  n.Rdisplace.dyads <- length(dyads.displace$actor)

  Rcrowd.density <- n.Rcrowd.dyads/n.total.dyads
  Rdisplace.density <- n.Rdisplace.dyads/n.total.dyads
  
  # Pool basic summaries
  Rcd.basic <- cbind.data.frame(run.code, run.cor, 
                                n.Rcrowd, n.Rdisplace,
                                prop.Rcrowd, prop.Rdisplace, 
                                n.Rcrowd.dyads, n.Rdisplace.dyads,
                                Rcrowd.density, Rdisplace.density
                                )
  #str(Rcd.basic)
  
  ## START NETWORK ANALYSES ----
  ref.modelW0s <- merge(dyad.list, run.data, all.x=TRUE,
                       by=c("actor", "subject")) #head(run.dataW0s)
  
  #print a check
  check <- length(ref.modelW0s$actor)
  print(check)
  
  #make runID character (not factor)
  ref.modelW0s$runID <- as.character(ref.modelW0s$runID)
 
   #fill newly-merged data with 0's where no interactions
  ref.modelW0s[is.na(ref.modelW0s)] <- 0
  
###### separate analyses by type (crowd vs displace)
  
  ### CROWD
  
  ref.crowd.mx <- reshape2::dcast(ref.modelW0s, actor~subject, value.var="random.crowd") #head(ref.crowd.mx)
  ref.crowd.mx[is.na(ref.crowd.mx)] <- 0 #for linearity measure, matrix needs to be fully filled, no NAs
  ref.crowd.mx <- matrix.please(ref.crowd.mx)
  
   # network graph
  graph.ref.crowd <- graph_from_adjacency_matrix(ref.crowd.mx, "directed", weighted=T, diag=F)
  
#individual level
  #out-degree
  Rcrowd.degree <- igraph::degree(graph.ref.crowd, mode = c("out"))
  
   # strength
   Rcrowd.strength <- strength(graph.ref.crowd, mode = c("out"), weights = E(graph.ref.crowd)$weight)
  
   # betweenness centrality
   Rcrowd.between <- igraph::betweenness(graph.ref.crowd, directed = TRUE,
                                    weights = E(graph.ref.crowd)$weight)
  
  # eigenvector centrality
  Rcrowd.eigen <- eigen_centrality(graph.ref.crowd, directed = TRUE, weights = E(graph.ref.crowd)$weight)
  Rcrowd.ec <- Rcrowd.eigen$vector
  Rcrowd.ec.value <- Rcrowd.eigen$value

#group-level  
  # average path length
  Rcrowd.apl <- mean_distance(graph.ref.crowd) #average.path.length(graph.ref.crowd) 
  
  # efficiency
  Rcrowd.ei <- effective_information(graph.ref.crowd, effectiveness = FALSE)
  Rcrowd.eff <- Rcrowd.ei/log2(length(bird.list))
  
  #Find proportion unknown relationships, a measure of sparseness
  Rcrowd.prunk <- EloRating::prunk(ref.crowd.mx)
  Rcrowd.prunk.pu <- as.numeric(Rcrowd.prunk[1])
  Rcrowd.prunk.dyads <- as.numeric(Rcrowd.prunk[2])
  
# dominance  
  #Find linearity
  lin.crowd<- EloRating::h.index(ref.crowd.mx, loops = 1000)
  Rcrowd.linearity <- lin.crowd[3,2]

  #Find steepness
  Rcrowd.steepness <- steepness::getStp(ref.crowd.mx, method="Dij")
  
  #Triangle transitivity
  Rcrowd.transi <- EloRating::transitivity(ref.crowd.mx, runs = 1000)
  Rcrowd.transi.Pt <- Rcrowd.transi[1]  # proportion of transitive triads
  Rcrowd.transi.ttri <- Rcrowd.transi[2]  # triangle transitivity
  
  
  ### POOL crowd hierarchy
  Rcrowd.hier <- cbind.data.frame(
                                  Rcrowd.degree,
                                  Rcrowd.strength,
                                  Rcrowd.between,
                                  Rcrowd.ec, Rcrowd.ec.value,
                                  Rcrowd.apl,
                                  Rcrowd.ei, Rcrowd.eff,
                                  Rcrowd.linearity, 
                                  Rcrowd.steepness, 
                                  Rcrowd.prunk.pu, Rcrowd.prunk.dyads,
                                  Rcrowd.transi.Pt, Rcrowd.transi.ttri
                                  )
  
  ### DISPLACEMENT
  
  ref.disp.mx <- reshape2::dcast(ref.modelW0s, actor~subject, value.var="random.displace") #head(ref.disp.mx)
  ref.disp.mx[is.na(ref.disp.mx)] <- 0
  ref.disp.mx <- matrix.please(ref.disp.mx)
  
  # network
  graph.ref.displace <- graph_from_adjacency_matrix(ref.disp.mx, "directed", weighted=T, diag=F)

# individual-based metrics    
  #out-degree
  Rdisp.degree <- igraph::degree(graph.ref.displace, mode = c("out"))
  
   # strength
   Rdisp.strength <- strength(graph.ref.displace, mode = c("out"), weights = E(graph.ref.displace)$weight)
  
   # betweenness centrality
   Rdisp.between <- igraph::betweenness(graph.ref.displace, directed = TRUE,
                                    weights = E(graph.ref.displace)$weight)
  
  # eigenvector centrality
  Rdisp.eigen <- eigen_centrality(graph.ref.displace, directed = TRUE, weights = E(graph.ref.displace)$weight)
  Rdisp.ec <- Rdisp.eigen$vector
  Rdisp.ec.value <- Rdisp.eigen$value
  
# group-level metrics
  # average path length
  Rdisp.apl <- mean_distance(graph.ref.displace) #average.path.length(graph.ref.disp) 
  
  # efficiency
  Rdisp.ei <- effective_information(graph.ref.displace, effectiveness = FALSE)
  Rdisp.eff <- Rdisp.ei/log2(length(bird.list))
  
  #Find proportion unknown relationships, a measure of sparseness
  Rdisp.prunk <- EloRating::prunk(ref.disp.mx)
  Rdisp.prunk.pu <- as.numeric(Rdisp.prunk[1])
  Rdisp.prunk.dyads <- as.numeric(Rdisp.prunk[2])
  
  
# dominance  
  #Find linearity
  lin.disp<- EloRating::h.index(ref.disp.mx, loops = 1000)
  Rdisp.linearity <- lin.disp[3,2]

  #Find steepness
  Rdisp.steepness <- steepness::getStp(ref.disp.mx, method="Dij")
  
  #Triangle transitivity
  Rdisp.transi <- EloRating::transitivity(ref.disp.mx, runs = 1000)
  Rdisp.transi.Pt <- Rdisp.transi[1]  # proportion of transitive triads
  Rdisp.transi.ttri <- Rdisp.transi[2]  # triangle transitivity
  

  #### POOL displace hierarchy
  Rdisp.hier <- cbind.data.frame( Rdisp.degree,
                                  Rdisp.strength,
                                  Rdisp.between,
                                  Rdisp.ec, Rdisp.ec.value,
                                  Rdisp.apl,
                                  Rdisp.ei, Rdisp.eff,
                                  Rdisp.linearity, 
                                  Rdisp.steepness, 
                                  Rdisp.prunk.pu, Rdisp.prunk.dyads,
                                  Rdisp.transi.Pt, Rdisp.transi.ttri
                                  )
  
  ### matrix correlation ----
  ref2.cor.result <- mantel(ref.crowd.mx, ref.disp.mx, method = "spearman") # vegan package
  ref2.cor.matrix <- ref2.cor.result$statistic
  
  ## correlation individual-based metric ----
  R.cor.strength <- cor(Rcrowd.strength, Rdisp.strength,  
               method = "spearman")
  R.cor.between <- cor(Rcrowd.between, Rdisp.between,  
               method = "spearman")
  R.cor.eigen <- cor(Rcrowd.ec, Rdisp.ec,  
               method = "spearman")
  
  # POOL ALL DATA ----
  run.sums <- cbind.data.frame(Rcd.basic, Rcrowd.hier, Rdisp.hier, ref2.cor.matrix, R.cor.strength, R.cor.between, R.cor.eigen )
  ref.model.summaries <- rbind(ref.model.summaries, run.sums)
}

end.time <- Sys.time()

#Time to run:
end.time - start.time

#t(ref.model.summaries)
head(ref.model.summaries)

# export data
write.csv(ref.model.summaries, "ANALYZE-SUMM.refmodel2.csv")
```


#### Loop to produce aggression strategies

Finding the strategies for displacements (takes awhile!). 
```{r}
head(ref.model)
#run.s #list of all run names
#head(ref.model)
#head(ANALYZE_refmodelCOR_CxD)

run.s <- as.character(unique(ref.model$runID))
n.total.dyads <- length(dyad.list$dyadID) #total number of dyads 

#make empty dataframe to fill
ref.model.strategies <- data.frame(run.code=character(), 
                                   type=character(), 
                                   fp.Rdisp=numeric(), 
                                   strategy.Rdisp=numeric()
                                  )

refmodel.blurdata <- data.frame(runID=character(),
                                type=character(),
                                blur=numeric(),
                                focus=numeric(),
                                focus_ci_hi=numeric(),
                                focus_ci_lo=numeric(),
                                position=numeric(),
                                position_ci_hi=numeric(),
                                position_ci_lo=numeric()
                                )
#run=2

#n.Odisplace <- sum(run.data$observed.displace)
start.time <- Sys.time()

for(run in seq_along(unique(ref.model$runID))){
  run.code <- run.s[run]
  print(run.code)
  run.data <- subset(ref.model, runID==run.code)

  ref.modelW0s <- merge(dyad.list, run.data, all.x=TRUE,
                       by=c("actor", "subject")) #head(run.dataW0s)
  
  #print a check
  check <- length(ref.modelW0s$actor)
  print(check)
  
  #make runID character (not factor)
  ref.modelW0s$runID <- as.character(ref.modelW0s$runID)
 
   #fill newly-merged data with 0's where no interactions
  ref.modelW0s[is.na(ref.modelW0s)] <- 0
  

  ### DISPLACEMENT
  
  ref.disp.mx <- reshape2::dcast(ref.modelW0s, actor~subject, value.var="random.displace") #head(ref.disp.mx)
  ref.disp.mx[is.na(ref.disp.mx)] <- 0
  ref.disp.mx <- matrix.please(ref.disp.mx)
  
  # Compute focus & position
  focus.Rdisp <- dom_focus(ref.disp.mx)
  position.Rdisp <- dom_position(ref.disp.mx)
  fp.Rdisp <- cbind.data.frame(focus.Rdisp, position.Rdisp)
  colnames(fp.Rdisp) <- c("focus", "position")

  #Compute blur models
  blur.Rdisp <- dom_make_blur_data(ref.disp.mx)

  #compile displace summary
  runID <- rep(run.code, length(blur.Rdisp$blur))
  type <- rep("displace", length(blur.Rdisp$blur))
  blur.displace <- cbind.data.frame(runID, type, blur.Rdisp)
  
  #Find strategy
  strategy.Rdisp <- dom_categorize_strategy(data=fp.Rdisp, blur_data=blur.Rdisp)
  
  
  ### POOL ALL DATA
  
  #pool blur data
  refmodel.blurdata <- rbind.data.frame(refmodel.blurdata, blur.displace)
  
  #pool strategy/run data
  run.strategy.sums <- cbind.data.frame(run.code, 
                                        focus.Rdisp, position.Rdisp, strategy.Rdisp
                                        )
  
  ref.model.strategies <- rbind(ref.model.strategies, run.strategy.sums)

  
}

end.time <- Sys.time()

#Time to run:
end.time - start.time

#t(ref.model.strategies)
head(ref.model.strategies)


# check right refmodel used
sum(ref.model$crowd[ref.model$runID=="run003"])
sum(ref.model$displace[ref.model$runID=="run050"])

# export data
write.csv(ref.model.strategies, file = "ANALYZE_SUMM.refmodel2.strategies.csv")	
write.csv(refmodel.blurdata, file = "ANALYZE_SUMM.refmodel2.blurs.csv")
```


## Correlation between crowds and displacements

How are displacements and crowds correlated in the random runs?
NOTE: the package plyr conflicts with the package dplyr causing some weird behavior. Be VERY CAREFUL if you are using both! You can use detach(package:plyr) to de-activate plyr later in a script

NOTE: we perform a correlation matrix when we summarize our reference model and in our paper

```{r}
head(ref.model)

require(plyr)
func <- function(ref.model)
{
return(data.frame(COR = cor(ref.model$displace, ref.model$crowd, method = "spearman")))
}

COR.refmodel <- ddply(ref.model, .(runID), func)

#find mean of reference model correlations
COR.refmodel.mean <- mean(COR.refmodel$COR)


#PLOT
par(mfrow=c(1,1))

run.s <- unique(ref.model$runID)

replicates <- 100

for(run in 1:replicates){
  run.code <- run.s[run]
  run.data <- subset(ref.model, runID==run.code)
  run.corr <- subset(COR.refmodel, runID==run.code)
  with (run.data, plot (displace, crowd, main=paste0(run.code, " Correlation=", round(run.corr$COR, 2))))
  with (run.data, abline(lm(crowd~displace), col="grey")) # regression line (y~x)
}

detach(package:plyr)

# WRITE DATA  for SAMPLED data
head(COR.refmodel)
write.csv(COR.refmodel, file = "ANALYZE_sparse_refmodelCOR.csv")	
```


