---
title: "Reference model 1"
author: "Annemarie"
date: "16/06/2020"
output: html_document
---

All the analyses for reference model 1, where we randomized number of interactions between interaction types. 

# Load packages
```{r, include=FALSE}
library(dplyr)
library(tidyr)
library(reshape2)
library(stringr)
library(purrr)
library(EloRating)
library(domstruc) #install_github("danm0nster/domstruc")
```

# Import data
```{r}
ANALYZE_per0_aggCD <- read.csv(file="ANALYZE-per0-aggCD.csv") %>% 
  rename(rowID=X) # only necessary if the first column is an X 
agg.data <- subset(ANALYZE_per0_aggDC, days=="last") 
observed <- agg.data

head(observed)
str(observed)

# check 
unique(observed$actor)
unique(observed$subject)
unique(observed$date)

sum(observed$crowd)
sum(observed$displace)
sum(observed$total.DC)

```


## Helper functions & data
```{r}

#make matrix function (from dataframe, 1st col with row names)
matrix.please<-function(x) {
  m<-as.matrix(x[,-1])
  rownames(m)<-x[,1]
  m
}

#list of all valid color combinations/bird IDs
bird.list <- c("BBB","BBG","BBR","BGB","BGG","BRB","BRR","GBB","GBG","GGB","GGG","GGR","GRG","GRR","RBB","RBG","RBR","RGG","RGR","RRG","RRR")
length(bird.list)

## dataframe of all possible dyad combinations
dyad.list <- expand.grid(bird.list, bird.list) #head(dyad.list)
names(dyad.list) <- c("actor", "subject")
dyad.list <- subset(dyad.list, actor!=subject)
dyad.list$dyadID <- paste(dyad.list$actor, dyad.list$subject, sep="-")
dyad.list$actor <- as.character(dyad.list$actor)
dyad.list$subject <- as.character(dyad.list$subject)
head(dyad.list)

```



# Summarize data
Find total events per dyad (pool across days)

```{r}
observedXdyad <- observed
observedXdyad <- observedXdyad %>% 
                  dplyr::group_by(actor, subject) %>% 
                  dplyr::summarise(observed.crowd=sum(crowd),
                          observed.displace=sum(displace),
                          observed.totalCD=sum(total.DC)) 
observedXdyad <- as.data.frame(observedXdyad)   
observedXdyad$actor <- as.character(observedXdyad$actor)
observedXdyad$subject <- as.character(observedXdyad$subject)
```




#### 1. Create reference model
Test real patterns with randomized reference model: Using this reference model, we test whether randomizing the number of interactions across two behavior types results in indistinguishable structural patterns in a reference model compared to the observed data.

Change: Randomly re-allocate aggression to number of displacements and number of crowds by dyad. 
Preserve: Keep total events (n disp + n crowd) by dyad the same

```{r}
# Produce reference model
head(observedXdyad) 

replicates <- 100

#run=2

#make empty dataframe to write loop results into
ref.model <- data.frame(actor=character(),
                       subject=character(),
                       observed.crowd=numeric(),
                       observed.displace=numeric(),
                       observed.totalCD=numeric(),
                       random.crowd=numeric(),
                       random.displace=numeric())


for (run in 1:replicates) {
  r.seed <- run
  RNGkind(sample.kind="default") # code to get same random numbers across operating systems (mac vs windows)
  set.seed(r.seed)
  ref.data <- observedXdyad %>%
                rowwise() %>%
                dplyr::mutate(random.displace = sample(0:observed.totalCD, 1),
                              random.crowd = observed.totalCD-random.displace)
  
  ref.data$runID <- rep(paste0("run",str_pad(r.seed, 3, side="left", pad = "0")),    length(observedXdyad$actor)) #generates run IDs that are all 3 digits (001-100), change to 4 digit-padding if running >999
    
  ref.data <- as.data.frame(ref.data)
  ref.model <- rbind.data.frame(ref.model, ref.data)
}


## check ref model output

#check that the randomized interactions vary from the observed interactions
ref.model$random.displace == ref.model$observed.displace
ref.model$random.crowd == ref.model$observed.crowd

#check that runs are different from each other
head(subset(ref.model, runID=="run001"))
head(subset(ref.model, runID=="run002"))
head(subset(ref.model, runID=="run003"))
head(subset(ref.model, runID=="run011"))

#check that all runs ran as expected
unique(ref.model$runID)
length(unique(ref.model$runID))


######
        #loop to compare random to observed displacements for each runID
uniqrun = unique(ref.model$runID)
randvobs = data.frame(nmatches = numeric(), compar = character())
nrowxrun = nrow(observedXdyad)

for(i in uniqrun){
        nmatches = sum(ref.model[ref.model$runID==i,]$random.displace==ref.model[ref.model$runID==i,]$observed.displace)
        run = i
        df = as.data.frame(nmatches, run)
        randvobs=rbind.data.frame(randvobs,df)
        
        if(sum(randvobs[,1]==nrowxrun)>0){
                                break
                        } #if there is at least 1 occurrence of a complete match between random.displace and observed.displace columns, then break the loop. No more comparisons will be made. Look at end of randvobs to see which runs had a complete match
}

#check
randvobs


        #loop that compares random.crowds from each run to all other runs 
uniqrun = unique(ref.model$runID)
empdf = data.frame(nmatches = numeric(), compar = character())
nrowxrun = nrow(observedXdyad)

start.time = Sys.time()
        for (i in uniqrun) {
                
                #make uniqrunj (a list of all runIDs that come after i)
                n= which(!is.na(match(uniqrun,i))) 
                uniqrunj = uniqrun[n:length(uniqrun)] 
                uniqrunj = uniqrunj[uniqrunj!=i] 
                        
                for (j in uniqrunj) {
                  
                        nmatches = sum(ref.model[ref.model$runID==i,]$random.crowd==ref.model[ref.model$runID==j,]$random.crowd)    # gives the number of dyads for which the number of crowds in run i matches that in run j
                        
                        compar = paste(i,j, sep="-") #a vector to keep track of which runs are being compared
                        df = as.data.frame(nmatches,compar)
                        empdf=rbind.data.frame(empdf,df)
                  
                        if(sum(empdf[,1]==nrowxrun)>0){
                                break
                        } #if there is at least 1 occurrence of a complete match between random.crowd columns from the two runs, then break the inner loop
                        
                }
        
                val= as.numeric(substr(i,4,6))
                if(val %% 10==0) {
              cat(paste0("iteration: ", val, "\n"))
                   }# see loop progress at every 10 iterations
                               
                if(sum(empdf[,1]==nrowxrun)>0){
                        break
                } #if there is at least 1 occurrence of a complete match between random.crowd columns from the two runs, then break the entire loop - no more comparisons will be made. Look at end of empdf to see which runs had a complete match
                
        } #returns empdf with number of matches for each pair of runs being compared
tail(empdf)

end.time = Sys.time()

#Time to run:
end.time - start.time

#rm(i,j, df, n, uniqrunj, val)
#i="run050"
#j="run007"

####

# WRITE DATA 
write.csv(ref.model, file = "ANALYZE-refmodeldata.csv")

```



Now, we can summarize our results using this reference model.

```{r}
# load the reference model data
ANALYZE_refmodeldata <- read.csv(file="ANALYZE-refmodeldata.csv")
head(ANALYZE_refmodeldata)

ref.model <- ANALYZE_refmodeldata 

```

#### 2. Correlation between crowds and displacements
How are displacements and crowds correlated in the random runs?
NOTE: the package plyr conflicts with the package dplyr causing some weird behavior. Be VERY CAREFUL if you are using both! You can use detach(package:plyr) to de-activate plyr later in a script

```{r}
head(ref.model)

require(plyr)
func <- function(ref.model)
{
return(data.frame(COR = cor(ref.model$random.displace, ref.model$random.crowd)))
}

COR.refmodel <- ddply(ref.model, .(runID), func)

#find observed correlation between number of crowds and number of displacements
COR.observed <- cor(ref.model$observed.crowd, ref.model$observed.displace)

#find mean of reference model correlations
COR.refmodel.mean <- mean(COR.refmodel$COR)


#PLOT
par(mfrow=c(1,1))

run.s <- unique(ref.model$runID)

replicates <- 100

for(run in 1:replicates){
  run.code <- run.s[run]
  run.data <- subset(ref.model, runID==run.code)
  run.corr <- subset(COR.refmodel, runID==run.code)
  with (run.data, plot (random.displace, random.crowd, main=paste0(run.code, " Correlation=", round(run.corr$COR, 2))))
  with (run.data, abline(lm(random.crowd~random.displace), col="grey")) # regression line (y~x)
}

# WRITE DATA
write.csv(COR.refmodel, file = "ANALYZE_refmodelCOR_CxD.csv")	

```


#### 3. summarize ref model

```{r}
# import data (only necessary when you only run this part)

ANALYZE_refmodeldata <- read.csv(file="ANALYZE-refmodeldata.csv")# reference model data
head(ANALYZE_refmodeldata)
ref.model <- ANALYZE_refmodeldata 

ANALYZE_refmodelCOR_CxD <- read.csv(file="ANALYZE_refmodelCOR_CxD.csv") %>% 
  select(-X) # correlation reference model
COR.refmodel <- ANALYZE_refmodelCOR_CxD

```

```{r}
# Summarize data: find total events per dyad (pool across days)
observedXdyad <- observed
observedXdyad <- observedXdyad %>% 
                  dplyr::group_by(actor, subject) %>% 
                  dplyr::summarise(observed.crowd=sum(crowd),
                          observed.displace=sum(displace),
                          observed.totalCD=sum(total.DC)) 
observedXdyad <- as.data.frame(observedXdyad)   
observedXdyad$actor <- as.character(observedXdyad$actor)
observedXdyad$subject <- as.character(observedXdyad$subject)

```

## Check observed data 
Need to make sure that all birds in the group are included in each network (fill no interactions with zeros)

```{r}


# list of all dyads
head(dyad.list)
str(dyad.list)

# observed aggression by dyad by behavior
head(observedXdyad)
str(observedXdyad)

#subset(observedXdyad, actor==subject) #check that we don't have any self-loops

# merge data with full dyad list in case any didn't interact
observedXdyad.alldyads <- merge(observedXdyad, dyad.list, 
                           by=c("actor", "subject"), 
                           all.y=TRUE)
head(observedXdyad.alldyads)
str(observedXdyad.alldyads) #should have 420 dyads -> length(dyad.list$dyadID)

# convert NA to 0's
observedXdyad.alldyads[is.na(observedXdyad.alldyads)] <- 0

#check
head(observedXdyad.alldyads)
length(observedXdyad.alldyads$dyadID) # with 21 birds, we should have 420 dyads total


# cast into square matrix with actor vertically in first column, all subjects horizontally in first row
OBS.crd.mx <- dcast(observedXdyad.alldyads, actor~subject, value.var="observed.crowd")

# check that actors and subjects are correct
subset(observedXdyad.alldyads, actor=="BBB" & subject=="GGG")
subset(observedXdyad.alldyads, actor=="GGG" & subject=="BBB")
OBS.crd.mx[1,] #BBB wins against GGG
OBS.crd.mx[11,] # GGG wins against BBB
```


# Loop to produce all reference model summaries (except strategies)
Strategies take about 12 min each to run (12x100 crowd, 12x100 disp = 40 hours!)
Loop other than strategies takes about 14 min for full dataset and 4 min for sampled dataset. 

```{r}
#run.s #list of all run names
#head(ref.model)
#head(COR.refmodel)

run.s <- as.character(unique(ref.model$runID))
n.total.dyads <- length(dyad.list$dyadID) #total number of dyads 
replicates <- 100 #how many runs of the reference model were done

#make empty dataframe to fill
ref.model.summaries <- data.frame(run.code=character(), 
                                  run.cor=numeric(), 
                                  n.Rcrowd=numeric(), 
                                  n.Rdisplace=numeric(),
                                  prop.Rcrowd=numeric(), 
                                  prop.Rdisplace=numeric(), 
                                  n.Rcrowd.dyads=numeric(), 
                                  n.Rdisplace.dyads=numeric(),
                                  Rcrowd.density=numeric(), 
                                  Rdisplace.density=numeric(),
                                  Rcrowd.linearity=numeric(), 
                                  Rcrowd.steepness=numeric(), 
                                  Rcrowd.prunk.pu=numeric(), 
                                  Rcrowd.prunk.dyads=numeric(),
                                  Rcrowd.transi.Pt=numeric(), 
                                  Rcrowd.transi.ttri=numeric(),
                                  Rdisp.linearity=numeric(), 
                                  Rdisp.steepness=numeric(), 
                                  Rdisp.prunk.pu=numeric(), 
                                  Rdisp.prunk.dyads=numeric(),
                                  Rdisp.transi.Pt=numeric(), 
                                  Rdisp.transi.ttri=numeric()
                                  )
#run=2

#n.Ocrowd <- sum(run.data$observed.crowd)
#n.Odisplace <- sum(run.data$observed.displace)
start.time <- Sys.time()

for(run in 1:replicates){
  run.code <- run.s[run]
  print(run.code)
  run.data <- subset(ref.model, runID==run.code)
  run.corr <- subset(COR.refmodel, runID==run.code)
  run.cor <- run.corr[1,2]
  
  #summarize basic info
  n.Rcrowd <- sum(run.data$random.crowd)
  n.Rdisplace <- sum(run.data$random.displace)
  prop.Rdisplace <- n.Rdisplace/(n.Rdisplace+n.Rcrowd)
  prop.Rcrowd <- n.Rcrowd/(n.Rdisplace+n.Rcrowd)

  dyads.crowd <- subset(run.data, random.crowd>0)
  dyads.displace <- subset(run.data, random.displace>0)

  n.Rcrowd.dyads <- length(dyads.crowd$actor)
  n.Rdisplace.dyads <- length(dyads.displace$actor)

  Rcrowd.density <- n.Rcrowd.dyads/n.total.dyads
  Rdisplace.density <- n.Rdisplace.dyads/n.total.dyads
  
  # Pool basic summaries
  Rcd.basic <- cbind.data.frame(run.code, run.cor, 
                                n.Rcrowd, n.Rdisplace,
                                prop.Rcrowd, prop.Rdisplace, 
                                n.Rcrowd.dyads, n.Rdisplace.dyads,
                                Rcrowd.density, Rdisplace.density
                                )
  #str(Rcd.basic)
  
  # START NETWORK ANALYSES
  ref.modelW0s <- merge(dyad.list, run.data, all.x=TRUE,
                       by=c("actor", "subject")) #head(run.dataW0s)
  
  #print a check
  check <- length(ref.modelW0s$actor)
  print(check)
  
  #make runID character (not factor)
  ref.modelW0s$runID <- as.character(ref.modelW0s$runID)
 
   #fill newly-merged data with 0's where no interactions
  ref.modelW0s[is.na(ref.modelW0s)] <- 0
  
###### separate analyses by type (crowd vs displace)
  
  ### CROWD
  
  ref.crowd.mx <- reshape2::dcast(ref.modelW0s, actor~subject, value.var="random.crowd") #head(ref.crowd.mx)
  ref.crowd.mx[is.na(ref.crowd.mx)] <- 0 #for linearity measure, matrix needs to be fully filled, no NAs
  ref.crowd.mx <- matrix.please(ref.crowd.mx)
  
  #Find linearity
  lin.crowd<- EloRating::h.index(ref.crowd.mx, loops = 1000)
  Rcrowd.linearity <- lin.crowd[3,2]

  #Find steepness
  Rcrowd.steepness <- steepness::getStp(ref.crowd.mx, method="Dij")
  
  #Find proportion unknown relationships, a measure of sparseness
  Rcrowd.prunk <- EloRating::prunk(ref.crowd.mx)
  Rcrowd.prunk.pu <- as.numeric(Rcrowd.prunk[1])
  Rcrowd.prunk.dyads <- as.numeric(Rcrowd.prunk[2])
  
  #Triangle transitivity
  Rcrowd.transi <- EloRating::transitivity(ref.crowd.mx, runs = 1000)
  Rcrowd.transi.Pt <- Rcrowd.transi[1]  # proportion of transitive triads
  Rcrowd.transi.ttri <- Rcrowd.transi[2]  # triangle transitivity
  
  
  ### POOL crowd hierarchy
  Rcrowd.hier <- cbind.data.frame(
                                  Rcrowd.linearity, 
                                  Rcrowd.steepness, 
                                  Rcrowd.prunk.pu, Rcrowd.prunk.dyads,
                                  Rcrowd.transi.Pt, Rcrowd.transi.ttri
                                  )
  
  ### DISPLACEMENT
  
  ref.disp.mx <- reshape2::dcast(ref.modelW0s, actor~subject, value.var="random.displace") #head(ref.disp.mx)
  ref.disp.mx[is.na(ref.disp.mx)] <- 0
  ref.disp.mx <- matrix.please(ref.disp.mx)
  
  #Find linearity
  lin.disp<- EloRating::h.index(ref.disp.mx, loops = 1000)
  Rdisp.linearity <- lin.disp[3,2]

  #Find steepness
  Rdisp.steepness <- steepness::getStp(ref.disp.mx, method="Dij")
  
  #Find proportion unknown relationships, a measure of sparseness
  Rdisp.prunk <- EloRating::prunk(ref.disp.mx)
  Rdisp.prunk.pu <- as.numeric(Rdisp.prunk[1])
  Rdisp.prunk.dyads <- as.numeric(Rdisp.prunk[2])
  
  #Triangle transitivity
  Rdisp.transi <- EloRating::transitivity(ref.disp.mx, runs = 1000)
  Rdisp.transi.Pt <- Rdisp.transi[1]  # proportion of transitive triads
  Rdisp.transi.ttri <- Rdisp.transi[2]  # triangle transitivity
  

  
  #### POOL displace hierarchy
  Rdisp.hier <- cbind.data.frame(
                                  Rdisp.linearity, 
                                  Rdisp.steepness, 
                                  Rdisp.prunk.pu, Rdisp.prunk.dyads,
                                  Rdisp.transi.Pt, Rdisp.transi.ttri
                                  )
  
  # POOL ALL DATA
  run.sums <- cbind.data.frame(Rcd.basic, Rcrowd.hier, Rdisp.hier)
  ref.model.summaries <- rbind(ref.model.summaries, run.sums)
}

end.time <- Sys.time()

#Time to run:
end.time - start.time

#t(ref.model.summaries)
head(ref.model.summaries)


# write data
write.csv(ref.model.summaries, "ANALYZE-SUMM.refmodel.csv")

```


#### 4. summarize ref model strategies

```{r}
# import data (only necessary when you only run this part)

ANALYZE_refmodeldata <- read.csv(file="ANALYZE-refmodeldata.csv") %>%# reference model data
 select(-X)
head(ANALYZE_refmodeldata)
ref.model <- ANALYZE_refmodeldata 
```

#Find strategies (takes awhile!)

```{r}
#run.s #list of all run names
#head(ref.model)

run.s <- as.character(unique(ref.model$runID))
n.total.dyads <- length(dyad.list$dyadID) 
replicates <- 100 #how many runs of the reference model were done

#make empty dataframe to fill
ref.model.strategies <- data.frame(run.code=character(), 
                                   type=character(), 
                                   fp.Rcrowd=numeric(), 
                                   strategy.Rcrowd=numeric(), 
                                   fp.Rdisp=numeric(), 
                                   strategy.Rdisp=numeric()
                                  )

refmodel.blurdata <- data.frame(runID=character(),
                                type=character(),
                                blur=numeric(),
                                focus=numeric(),
                                focus_ci_hi=numeric(),
                                focus_ci_lo=numeric(),
                                position=numeric(),
                                position_ci_hi=numeric(),
                                position_ci_lo=numeric()
                                )
#run=2

#n.Ocrowd <- sum(run.data$observed.crowd)
#n.Odisplace <- sum(run.data$observed.displace)
start.time <- Sys.time()

for(run in 1:replicates){
  run.code <- run.s[run]
  print(run.code)
  run.data <- subset(ref.model, runID==run.code)

  ref.modelW0s <- merge(dyad.list, run.data, all.x=TRUE,
                       by=c("actor", "subject")) #head(run.dataW0s)
  
  #print a check
  check <- length(ref.modelW0s$actor)
  print(check)
  
  #make runID character (not factor)
  ref.modelW0s$runID <- as.character(ref.modelW0s$runID)
 
   #fill newly-merged data with 0's where no interactions
  ref.modelW0s[is.na(ref.modelW0s)] <- 0
  
  ### CROWD
  
  ref.crowd.mx <- reshape2::dcast(ref.modelW0s, actor~subject, value.var="random.crowd") #head(ref.crowd.mx)
  ref.crowd.mx[is.na(ref.crowd.mx)] <- 0 #for linearity measure, matrix needs to be fully filled, no NAs
  ref.crowd.mx <- matrix.please(ref.crowd.mx)
  
  #CROWDS
  # Compute focus & position
  focus.Rcrowd <- dom_focus(ref.crowd.mx)
  position.Rcrowd <- dom_position(ref.crowd.mx)
  fp.Rcrowd <- cbind.data.frame(focus.Rcrowd, position.Rcrowd)
  colnames(fp.Rcrowd) <- c("focus", "position")
  
  #Compute blur models
  blur.Rcrowd <- dom_make_blur_data(ref.crowd.mx)
  
  #compile crowd summary
  runID <- rep(run.code, length(blur.Rcrowd$blur))
  type <- rep("crowd", length(blur.Rcrowd$blur))
  blur.crowd <- cbind.data.frame(runID, type, blur.Rcrowd)
  
  #Find strategy
  strategy.Rcrowd <- dom_categorize_strategy(data=fp.Rcrowd, blur_data=blur.Rcrowd)
  
  
  ### DISPLACEMENT
  
  ref.disp.mx <- reshape2::dcast(ref.modelW0s, actor~subject, value.var="random.displace") #head(ref.disp.mx)
  ref.disp.mx[is.na(ref.disp.mx)] <- 0
  ref.disp.mx <- matrix.please(ref.disp.mx)
  
  # DISPLACEMENTS
  # Compute focus & position
  focus.Rdisp <- dom_focus(ref.disp.mx)
  position.Rdisp <- dom_position(ref.disp.mx)
  fp.Rdisp <- cbind.data.frame(focus.Rdisp, position.Rdisp)
  colnames(fp.Rdisp) <- c("focus", "position")

  #Compute blur models
  blur.Rdisp <- dom_make_blur_data(ref.disp.mx)

  #compile displace summary
  runID <- rep(run.code, length(blur.Rdisp$blur))
  type <- rep("displace", length(blur.Rdisp$blur))
  blur.displace <- cbind.data.frame(runID, type, blur.Rdisp)
  
  #Find strategy
  strategy.Rdisp <- dom_categorize_strategy(data=fp.Rdisp, blur_data=blur.Rdisp)
  
  
  ### POOL ALL DATA
  
  #pool blur data
  refmodel.blurdata <- rbind.data.frame(refmodel.blurdata, blur.crowd, blur.displace)
  
  #pool strategy/run data
  run.strategy.sums <- cbind.data.frame(run.code, 
                                        focus.Rcrowd, position.Rcrowd,
                                        strategy.Rcrowd, 
                                        focus.Rdisp, position.Rdisp, strategy.Rdisp
                                        )
  
  ref.model.strategies <- rbind(ref.model.strategies, run.strategy.sums)

  
}

end.time <- Sys.time()

#Time to run:
end.time - start.time

#t(ref.model.strategies)
head(ref.model.strategies)


#WRITE DATA
write.csv(ref.model.strategies, file = "ANALYZE_SUMM.refmodel.strategies.csv", sep="")	write.csv(ref.model.blurdata, file = "ANALYZE_SUMM.refmodel.blurs.csv", sep="")
```




