---
title: "Visualizing results"
author: "Annemarie"
date: "01/06/2020"
output: html_document
---

Comparison of real vs randomized data in 4-6 plots:
- correlation crowds vs displacements
- density
- linearity
- steepness
- transitivity
- sparseness

# load packages
```{r setup, include=FALSE}
library(beanplot)
citation("beanplot")
```

# Import data

```{r}
ref.model.summaries<- read.csv("EXPLORE-SUMM.refmodel.csv")
observed.summary <- read.csv("EXPLORE-SUMM.observed.csv")
```

## Including Plots
# correlation
```{r pressure, echo=FALSE}
# correlation
beanplot(refmodel.summary$Rcor,  
         main="Distribution of correlations", 
         ylab="Correlation strength", 
         ylim=c(0, 1),
         las=1, col = alpha("black", 0.6)) 
points(1,observed.summary$obs.cor, col="#fc8d62", pch=7)

```
I added correlation to the refmodel.summaries dataset in our script. if we decide not to do that we have to load the correlation dataframe separately: 
```{r}
#reference model correlations of crowd vs displacements
EXPLORE_refmodelCOR_CxD <- read.csv(file="EXPLORE_refmodelCOR_CxD.csv")
correlationsCxD<-EXPLORE_refmodelCOR_CxD 
head(correlationsCxD)

beanplot(correlationsCxD$COR,  
         main="Distribution of correlations", 
         ylab="Correlation strength", 
         ylim=c(0, 1),
         las=1, col = alpha("black", 0.6)) 
points(1,observed.summary$obs.cor, col="#fc8d62", pch=7)

```




# Proportion of number of interactions by total observed
```{r, echo=FALSE}
# histogram dyadic relationships
beanplot(refmodel.summary$prop.Rcrowd, refmodel.summary$prop.Rdisplace,  
         main="Distribution of proportions", 
         ylab="# interactions/total observed", 
         ylim = c(min(observed.summary$prop.Obcrowd-0.1), 
                  max(observed.summary$prop.Obdisplace+0.1)),
         xlab = "interaction", names = c("crowds", "displacements"),
         col = alpha("black", 0.6))
points(1,observed.summary$prop.Obcrowd, col="#fc8d62", pch=7) # crowds
points(2,observed.summary$prop.Obdisplace, col="#8da0cb", pch=7) # crowds
```

# Density
```{r, echo=FALSE}
# density
beanplot(refmodel.summary$Rcrowd.density, refmodel.summary$Rdisplace.density,  
         main="Distribution of density", 
         ylab="prop agonistic dyads/total dyads", 
         ylim = c(min(observed.summary$Obcrowd.density-0.1), 
                  max(observed.summary$Obdisplace.density+0.1)),
         xlab = "interaction", names = c("crowds", "displacements"),
         col = alpha("black", 0.6))
points(1,observed.summary$Obcrowd.density, col="#fc8d62", pch=7) # crowds
points(2,observed.summary$Obdisplace.density, col="#8da0cb", pch=7) # displacements
```
# Linearity
```{r, echo=FALSE}
# linearity
beanplot(refmodel.summary$Rcrowd.linearity, refmodel.summary$Rdisp.linearity, 
         main="Distribution of linearity", 
         ylab="h index", log = "",
         ylim = c(min(observed.summary$Obcrowd.linearity-0.1), 
                  max(observed.summary$Obdisp.linearity+0.1)),
         xlab = "interaction", names = c("crowds", "displacements"),
         col = alpha("black", 0.6))
points(1,observed.summary$Obcrowd.linearity, col="#fc8d62", pch=7) # crowds
points(2,observed.summary$Obdisp.linearity, col="#8da0cb", pch=7) # displacements
```

# Steepness
```{r, echo=FALSE}
beanplot(refmodel.summary$Rcrowd.steepness, refmodel.summary$Rdisp.steepness,
         main="Distribution of steepness", 
         ylab="Dij", las=1, log = "",
         ylim = c(min(observed.summary$Obcrowd.steepness-0.1), 
        max(refmodel.summary$Rdisp.steepness,observed.summary$Obdisp.steepness+0.1)),
         xlab = "interaction", names = c("crowds", "displacements"),
         col = alpha("black", 0.6))
points(1,observed.summary$Obcrowd.steepness, col="#fc8d62", pch=7) # crowds
points(2,observed.summary$Obdisp.steepness, col="#8da0cb", pch=7) # displacements
```

# Sparseness
```{r}
# sparseness
beanplot(refmodel.summary$Rcrowd.prunk.pu, refmodel.summary$Rdisp.prunk.pu,
         main ="Distribution of sparseness", 
         ylab="Sparseness", 
         ylim=c(min(observed.summary$Obdisp.prunk.pu-0.1), 
                max(refmodel.summary$Rcrowd.prunk.pu,observed.summary$Obcrowd.prunk.pu+0.1)),
         las=1, xlab = "interaction", names = c("crowds", "displacements"),
         col = alpha("black", 0.6))
points(1,observed.summary$Obcrowd.prunk.pu, col="#fc8d62", pch=7) # crowds
points(2,observed.summary$Obdisp.prunk.pu, col="#8da0cb", pch=7) # displacements
```



# Transitivity
```{r}
# transitivity
beanplot(refmodel.summary$Rcrowd.transi.Pt, refmodel.summary$Rdisp.transi.Pt,
         main ="Distribution of transitivity", 
         ylab="Transitivity",
         ylim=c(min(observed.summary$Obcrowd.transi.Pt-0.1), 
                max(refmodel.summary$Rdisp.transi.Pt,observed.summary$Obdisp.transi.Pt+0.1)),
         las=1, xlab = "interaction", names = c("crowds", "displacements"),
         col = alpha("black", 0.6))
points(1,observed.summary$Obcrowd.transi.Pt, col="#fc8d62", pch=7) # crowds
points(2,observed.summary$Obdisp.transi.Pt, col="#8da0cb", pch=7) # displacements

```

# Combine plots
using par(mfrow=c(3,2)) for the plots with crowds and displacements (6 total), then correlation plot separate
```{r}

```

# sessioninfo
```{r}
sessionInfo()
```

